Cluster Difference Imaging Photometric Survey. III. An Open Cluster Spanning 500\,Parsecs

### 2021-03-25 cluster_paper/ms.tex
update ms

\newcommand{\kms}{\,km\,s$^{-1}$}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{---}
\revised{---}
\accepted{---}
\submitjournal{ApJL.}
\shorttitle{Core \& Halo of NGC\,2516}

@@

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be tidal escapees from the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  sensitive clustering algorithms.
  In this study, we focus on a halo with a tip-to-tip length of
  $\approx$500\,pc reported around the $\approx$150\,Myr open cluster
  NGC\,2516.
  The classical tidal radius of the cluster is $\lesssim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two thirds of kinematically selected halo members out to 250\,pc from the
  cluster center have rotation periods consistent with a
  gyrochronological age of 150\,Myr.
  A comparison sample of field stars shows no such trend.
  In addition, the lithium abundances of stars in the halo are
  elevated relative to the field, and are correlated with the stellar
  rotation rate in a manner consistent with previous observations of the
  Pleiades and Psc Eri.
  This work expands the set of confirmed NGC 2516 members by a factor
  of two, and is consistent with a paradigm in which the halos of many
  open clusters are more populous than their cores.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f11.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Position and orbit of NGC\,2516 in the Galaxy.} 
  Points denote halo members with $\pi/\sigma_\pi>20$.  In our
  galactic coordinate system, the Sun is at $\{X,Y,Z\} = \{ -8122, 0,
  +20.8\}\,{\rm pc}$.  The gray line traces the past and future
  cluster orbit.  In different projections, the line spans different
  time intervals: the longest time window is visible in the $(Z,X)$
  projection.  The large blue arrows denote the median cluster
  velocity after subtracting the local standard of rest: $\{v_{\rm X},
  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.  The sizes of
  the blue arrows are internally proportionate.  By comparison,
  $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms.
  \label{fig:XYZ}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{f8a.pdf}
			\includegraphics[width=0.5\textwidth]{f8b.pdf}
		}
	%		\includegraphics[width=0.7\textwidth]{f8a.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf Members of NGC\,2516 exist hundreds of parsecs from the core,
  and up to $\approx$5\kms\ from the core in tangential velocity.}
  Halo members
  that meet both kinematic and rotational membership criteria are shown in
  blue.  Halo members for which rotation periods should have been detectable,
  but were not detected, are shown in orange.  Projections are shown in galactic coordinates and
  (projection-corrected) 2D-tangential velocity space.  The histogram
  on the right bins versus 3D separation from the core and
  2D velocity difference. Bins have width 25\,pc
  in position, and 1.0\,km\,s$^{-1}$ in velocity.  {\bf Uncertainties
  in the fraction are taken as the quadrature sum of the rel unc of
  the numerator and denominator (FIXME)}
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

To quantify this, we compute the separation of the positions and 2D
velocities from the median \citetalias{cantatgaudin_gaia_2018} core
member.  The result is shown on the right side of
Figure~\ref{fig:physical_x_rotn}.  The halo extends to separations of
about 200\,pc in physical space from the cluster core.  This
corresponds to total a length of $\approx$500\,pc, depending on which
members of the halo are chosen as the ``tips'' on either end.  This
agrees broadly speaking with the structure of the halo reported by
\citetalias{kounkel_untangling_2019}.

@@

Figure~\ref{fig:lithiumcorehalo} shows the concatenation of the
Gaia-ESO and GALAH results, with the points colored by whether the
stars are in the core or the halo of the cluster.  The GALAH spectra
span a color range of $0<(Bp-Rp)_0<1.5$, due to the $V<14$ brightness
cutoff of the survey.  The overall depletion of stars redder than
$(Bp-Rp)_0\gtrsim1.4$, {\it i.e.}, later than K4.5V (0.71\,$M_\odot$),
is visible relative to the earlier K dwarfs.  For comparison, we have
also plotted the Li EWs measured by \citet{berger_identifying_2018}
for planet-hosting stars in the Kepler field (we interpolated from
their effective temperatures using the \citet{pecaut_mamajek_2013}
table\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version 2019.3.22.}.
The field star distribution peaks at late F-type stars.
The majority of kinematically selected stars in the core and the halo
show lithium equivalent widths substantially in excess of these field
stars.

We can go one step further, and match our sample of kinematically
selected stars with spectra against the TESS rotators.
The result is shown in Figure~\ref{fig:lithiumrot}
At fixed stellar mass
(and age), the rapid rotators tend to show elevated lithium equivalent
widths. This effect is mostly apparent in the K dwarfs.  Similar
trends in the Pleiades were noted nearly three decades ago by
\citet{soderblom_evolution_1993}.  More recent studies have found
comparable results in the Pleiades, the Psc-Eri stream, and M\,35
\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.

% TODO: this paragraph more or less parrots the processes described by
% Bouvier+20. are there others? is there a better way to describe
% these?

An important point regarding the lithium-rotation trend is that it has
been shown to carry over beyond just equivalent widths, and to the
actual abundance of lithium in the photosphere, A(Li).  A priori, one
might worry that the trend is tautological: perhaps some feature the
stellar equatorial velocity, or the microturbulent velocity at the
stellar surface, are larger for faster rotators and this could affect
the formation conditions of the lithium doublet.
%FIXME improve these words
Soderblom+1993 calculated the relevant LTE curves of growth for the
lithium line, and argued that this could not be the case.
Since then, LindAsplund+2009 have calculated the relevant non-LTE
correction, and has shown that it affects the abundances by
$<0.5\,{\rm dex}$.
The order of magnitude of the rotation-abundance trend is 2-3\,dex.

%     * OK, then to convert to an ABUNDANCE, follow Bouvier+2018. In other words,
%       just interpolate against the LTE curve of growth tables from
%       Soderblom+1993.
% 
%     * Nb. the Non-LTE correction has been shown to be <0.5 dex (but
%       >0.1 dex, for sun-like stars with Li abundances above 2 dex,
%       >which is the regime we're interested in for Pleaides and NGC
%       >2516).
%       CITE: Lind, Asplund, + 2009
%       https://ui.adsabs.harvard.edu/abs/2009A%26A...503..541L/abstract

Processes both internal and external to the star have been suggested to explain
the lithium-rotation trend (CITE see the recent review by Bouvier+20).
One explanation based in the stellar interior could be that the
convective mixing efficiency is anticorrelated with the surface
rotation (e.g., CITE Siess + Livio 1997, Baraffe+2017).  Another
possibility could be that stronger magnetic fields in the star's
interior inhibit convection (READ e.g., Ventura+98, Chabrier+07,
Somers + Pinsonneault 2014).  An external process that might also be
important is the effect of star-disk magnetic locking during the PMS
phase (CITE: magnetic braking).  Longer disk lifetimes would lead to
the star's outer convective zone being locked for longer while the
radiative core contracts.  The resulting differential rotation and
rotational mixing could drive the lithium depletion (CITE: Bouvier 08,
Eggenberger+12).
We have no preference between these possibilities---we simply note
that the \cn\ sample could be another helpful data point in
distinguishing them.

@@

@@

\subsection{What is the contamination fraction in the halo? Does it change vs. location?}

\subsection{Are the ``very slow rotators'' bonafide members?}
Unsure. They are not isotropically distributed around the
cluster... so either triggered star formation (CITE, CITE), or they're
actually field stars.

% \subsection{Mass differences between center and outer reaches?}

\subsection{Fast and blue rotators: are we going faster than other clusters?}

\subsection{Absolute age of NGC 2516}
Cummings \& Kalirai (2018) favor 150 Myr based on MSTO.
Godoy-Rivera+2021 favor 150 Myr... based on ...? what?
"""
Similar to Cummings et al. (2016), Cummings \& Kalirai (2018),
and Cummings et al. (2018) (see also Fritzewski et al. 2020 and Healy
\&
McCullough 2020)
"""

\subsection{Ways of doing this at different ages}
Some final discussion is warranted on the applicability of our
approach more generally.  Given a kinematically
identified population, this study uses photometry (Gaia HR diagrams;
TESS rotation) and spectroscopy (GALAH and Gaia-ESO lithium) to
confirm youth in the stars.  Other approaches are also possible.

One can use photometric criteria (e.g., Gaia HR diagram of
everything within some distance; Zari+18), or spectroscopic
criteria (CITE Berger+18), or with combinations thereof (e.g., CITE
Zerjal 2019, 2020, Zhou's work).  
It might even be possible to start using chemical abundance signatures
(CITE, GALAHDR3?).

These paths might in certain cases succeed in identifying more
complete samples of dispersed members of young populations, since they
do not require kinematic proximity.

@@

The combination of Gaia, TESS, and the GALAH and Gaia-ESO
spectroscopic surveys has clarified a few 
things about \cn.
\begin{itemize}
	\item {\it Overdensities observed by Gaia in five-dimensional (3-D position + 2-D velocity)
		phase space
		revealed a halo of stars spanning 500 parsecs in the plane of the galaxy.}
	The earlier reference to the full structure that we can find is by CITE KOUNKEL.
	The halo is more precisely described as a leading and trailing tail
	(Figure~\ref{fig:XYZ}), with the leading end on the interior of the
	cluster's
	galactic orbit.
	%
	\item {\it Isochronal, rotational, and lithium age-dating show that the halo is real and coeval with the core of \cn.}
    The Gaia EDR3 photometry and astrometry show a main-sequence turnoff consistent with
    an isochronal age of 150\,Myr for both the core and halo.
    The faintest M dwarfs in the core and halo are also elevated relative to a field star comparison sample, but using
    this to isochronally measure the age is challenging due to uncertainties in M dwarf atmospheric opacities.
    Gyrochronally, FIXME FIXME FIXME
    %
    \item {\it The identified halo extends to $\approx$250\,pc in separate from the cluster center, and out to $\approx$5\kms\ in velocity separation.}
    Figure~\ref{fig:physical_x_rotn} shows the overlap between kinematic
    and rotational cluster members. 
    %
    \item {\it The field star contamination rate increases at larger physical and 
    velocity separations from the cluster core.}
	Figure~\ref{fig:physical_x_rotn}

    The stars further from the c
    As a reminder, a star with a 1\kms\ velocity difference moves 1\,pc per Myr.

	More specifically, ``two thirds of kinematically selected halo members out to 250 pc from the cluster center have rotation periods consistent with a gyrochronological age of 150 Myr''.

	\item {\it Differential rotation in the Milky Way likely plays an important role in shaping the halos of nearby open clusters.}
	The Milky Way's differential rotation rate has a comparable
	order of magnitude ($\approx$-0.2\,deg\,Myr$^{-1}$\,kpc$^{-1}$) to that measured across the halo \citep{meingast_2021}.
	It therefore likely plays an important role in understanding the halo's formation.

\end{itemize}

>>>>

\newcommand{\kms}{\,km\,s$^{-1}$}
\newcommand{\ms}{\,m\,s$^{-1}$}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2.
%% Add "Submitted to " argument.
\received{---}
\revised{---}
\accepted{---}
\submitjournal{AAS journals.}
\shorttitle{Core \& Halo of NGC\,2516}

@@

% NB. the "classical tidal radius" is from e.g., Piskunov+08 or
% Kharchenko+2013. The latter givs a radius r2 of 0.71 degrees. At 380
% pc, corresponds to 4.7pc. The King tidal radius given by K13 is
% 7.70pc.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters. 
  The stars in these diffuse halos could be tidal escapees from the
  cluster cores, or they could be remnants of an initially dispersed
  star formation complex.
  They could also be field stars---the false positives of overly
  sensitive clustering algorithms.
  In this study, we focus on a halo with a tip-to-tip length of
  $\approx$500\,pc reported around the $\approx$150\,Myr open cluster
  NGC\,2516.
  The classical tidal radius of the cluster is $\lesssim$10\,pc.
  Combining photometry from Gaia, rotation periods from TESS, and
  lithium measurements from Gaia-ESO and GALAH, we find that the halo
  of NGC\,2516 is real and coeval with the core.
  Two thirds of kinematically selected halo members out to 250\,pc from the
  cluster center have rotation periods consistent with a
  gyrochronological age of 150\,Myr.
  A comparison sample of field stars shows no such trend.
  In addition, the lithium abundances of stars in the halo are
  elevated relative to the field, and are correlated with the stellar
  rotation rate in a manner consistent with previous observations.
  This work expands the set of confirmed NGC 2516 members by a factor
  of two, and is consistent with a paradigm in which the halos of many
  open clusters are more populous than their cores.
  We highlight the implications for spectroscopic survey targeting,
  open cluster dispersal, and planet searches around young stars.
\end{abstract}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.95\textwidth]{f11.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Position and orbit of NGC\,2516 in the Galaxy.} 
  Points denote halo members with $\pi/\sigma_\pi>20$.  In our
  galactic coordinate system, the Sun is at $\{X,Y,Z\} = \{ -8122, 0,
  +20.8\}\,{\rm pc}$.  The gray line traces the past and future
  cluster orbit.  In different projections, the line spans different
  time intervals: the longest time window is visible in the $(Z,X)$
  projection.  The large blue arrows denote the median cluster
  velocity after subtracting the local standard of rest: $\{v_{\rm X},
  v_{\rm Y}, v_{\rm Z}\} = \{-22.2, -25.3, -4.6\}$\kms.   By comparison,
  $\vec{v}_{\rm LSR} = \{12.9, 245.6, 7.78\}$\kms; in the inertial
  frame of the Galaxy, \cn\ is moving mostly in the $\hat{Y}$
  direction.
  The sizes of the blue arrows are internally proportionate. 
  \label{fig:XYZ}
	}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{f8a.pdf}
			\includegraphics[width=0.5\textwidth]{f8b.pdf}
		}
	%		\includegraphics[width=0.7\textwidth]{f8a.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ 
  {\bf Members of \cn\ exist hundreds of parsecs from the core,
  and up to $\approx$5\kms\ from the core in tangential velocity.}
  Halo members that meet both kinematic and rotational membership
  criteria (``Subset~$\mathcal{A}$'') are shown in blue.  Halo members for which rotation periods
  should have been detectable, but were not detected, are shown in
  orange.  
  The plotted stars are restricted to $0.5<(Bp-Rp)_0<1.2$.
  The {\it left panels} show galactic coordinates and
  2D-tangential velocity space, after
  correcting for projection effects
  (Appendix~\ref{app:vproj}).
  The histogram on the right bins versus 3D separation from the core
  and 2D velocity difference.  Bins have width 25\,pc in position, and
  1.0\,km\,s$^{-1}$ in velocity.  {\bf Uncertainties in the fraction
  are taken as the quadrature sum of the rel unc of the numerator and
  denominator (FIXME)}
  \label{fig:physical_x_rotn}
	}
\end{figure*}

@@

% some wide separation pairs and dist_pc:
% 5263077805266480000  5506439929389838592  421.083675
% 5506623616550497792  5273740456715689088  427.053947
% 5506439929389838592  5213536216500903680  427.557268
% 5273740456715689088  5488984598182200704  428.952934
% 5506439929389838592  5273740456715689088  438.906581

To quantify this, we compute the separation of the positions and 2D
velocities from the median \citetalias{cantatgaudin_gaia_2018} core
member.  The result is shown on the right side of
Figure~\ref{fig:physical_x_rotn}.  The halo extends to separations of
about 200\,pc in physical space from the cluster core.  The most
widely separated F2V-K2V stars with rotation periods (blue points in
Figure~\ref{fig:physical_x_rotn}) are separated by $\approx$430\,pc.
The total length of the structure is therefore 400--500\,pc, depending
on which members of the halo are chosen as the ``tips'' on either end.
Broadly speaking, this agrees with the structure of the halo reported
by \citetalias{kounkel_untangling_2019}.

@@

Figure~\ref{fig:lithiumcorehalo} shows the concatenation of the
Gaia-ESO and GALAH results, with the points colored by whether the
stars are in the core or the halo of the cluster.  The GALAH spectra
span a color range of $0<(Bp-Rp)_0<1.5$, due to the $V<14$ brightness
cutoff of the survey.  The overall increase of Li EW from
$0<(Bp-Rp)_0<1.0$ is thought to be driven by the temperatures in the
stellar chromospheres: hotter stars fully ionize Li out of its ground
state \citep[{\it e.g.},Figure~4 of][]{soderblom_evolution_1993}.  The
depletion of stars redder than $(Bp-Rp)_0\gtrsim1.4$, {\it i.e.},
later than K4.5V (0.71\,$M_\odot$), is visible relative to the earlier
K dwarfs.  For comparison, we have also plotted the Li EWs measured by
\citet{berger_identifying_2018} for planet-hosting stars in the Kepler
field (we interpolated from their effective temperatures using the
\citet{pecaut_mamajek_2013}
table\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt},
version 2019.3.22.}.  The field star distribution peaks at late F-type
stars.  The majority of kinematically selected stars in the core and
the halo show lithium equivalent widths substantially in excess of
these field stars.

We can go one step further, and match our sample of kinematically
selected stars with spectra against the TESS rotators.  The result is
shown in Figure~\ref{fig:lithiumrot}. At fixed stellar mass, the rapid
rotators have lithium equivalent widths an order of magnitude larger
than the slow rotators.  This effect is mostly apparent in the K
dwarfs.  Similar trends were noted in the Pleiades over three decades
ago \citep{butler_pleiades_1987}, and are most likely caused by
intrinsic differences in lithium abundance
\citep{soderblom_evolution_1993}.  Alternative explanations, such as
differences in line-formation conditions ({\it e.g.,} chromospheric
temperatures, microturbulent velocities, or the presence of starspots)
are incompatible with the available data.  The lithium-rotation
correlation is discussed further below
(Section~\ref{discussion:lithium}).

@@

%\subsection{Contamination fraction across the halo}
%\subsection{How did the halo form?}
%\subsection{Rotation in NGC\,2516}
%\subsection{The lithium-rotation correlation}
%\subsection{Age of NGC\,2516}

\subsection{Contamination fraction across the halo}

@@

\subsection{Rotation in NGC\,2516}

Gotta cite Fritzewski+2020.  (Maybe even do the Blanco1, Pleiades,
NGC2516, PscEri concatenation?)

  \paragraph{Slow sequence and fast sequence}
  %Why is there a slow and fast sequence?
  Slow: Skumanich spindown, Weber and Davis magnetic braking.
  This doesn't explain everything.
  Core envelope decoupling handwaves away the fast rotators? (Irwin,
  Aigrain, Monitor).
  Another fast rotator explanation: Naoz's merging of giant planets
  (could explain some fraction of the fast rotators -- not all of them).

  \paragraph{Stalling}
  Curtis19 found NGC6811 was stalling in the K dwarfs, i.e., Skumanich
  spindown is not monotonic in time.
  We're seeing latest K dwarfs spin slowest, earlier FGKs don't show
  much change.

  \paragraph{Are the ``very slow rotators'' bonafide members?}
  Unsure. They are not isotropically distributed around the
  cluster... so either triggered star formation (CITE, CITE), or they're
  actually field stars.

\subsection{The lithium-rotation correlation}
\label{discussion:lithium}

An important point regarding the lithium-rotation trend is that it has
been shown to carry over beyond just equivalent widths, and to the
actual abundance of lithium in the photosphere, A(Li).
\citet{soderblom_evolution_1993} calculated the relevant LTE curves of
growth for the Li\,6708\,\AA\ doublet, which can in principle be used to
convert our measured equivalent widths to abundances.  Non-LTE
corrections affect the abundances by $<0.5\,{\rm dex}$ and do not
siginificantly affect the global relation between Li abundance and
stellar temperature or rotation
\citep{carlsson_1994,lind_departures_2009}.

More recent studies of lithium and rotation since 
\citet{soderblom_evolution_1993} have expanded these results in the
Pleiades, the Psc-Eri stream, and M\,35, among other clusters
\citep{bouvier_pleiades_lirot_2018,arancibia_2020,jeffries_m35_li_2020}.
In \cn\ specifically, \citet{jeffries_rotation_1998} observed a
similar trend that that seen in Figure~\ref{fig:lithiumrot} when
analyzing 24 stars in the core of \cn.  Our concatenation of the
\citet{randich_gaiaeso_2018} Gaia-ESO results with the GALAH-DR3
spectra represent a significant expanasion in volume and color range
since these earlier studies.

What causes the correlation between lithium abundance and rotation at
fixed stellar mass?
{\bf FIXME: improve this paragraph}
%FIXME IMPROVE THIS PARAGRAPH.
Processes both internal and external to the star have been suggested to explain
the lithium-rotation trend (CITE see the recent review by Bouvier+20).
One explanation based in the stellar interior could be that the
convective mixing efficiency is anticorrelated with the surface
rotation (e.g., CITE Siess + Livio 1997, Baraffe+2017).  Another
possibility could be that stronger magnetic fields in the star's
interior inhibit convection (READ e.g., Ventura+98, Chabrier+07,
Somers + Pinsonneault 2014).  An external process that might also be
important is the effect of star-disk magnetic locking during the PMS
phase (CITE: magnetic braking).  Longer disk lifetimes would lead to
the star's outer convective zone being locked for longer while the
radiative core contracts.  The resulting differential rotation and
rotational mixing could drive the lithium depletion (CITE: Bouvier 08,
Eggenberger+12).
We have no preference between these possibilities---we simply note
that the \cn\ sample could be another helpful data point in
distinguishing them.

\subsection{Age of NGC\,2516}
Cummings \& Kalirai (2018) favor 150 Myr based on MSTO.
Godoy-Rivera+2021 favor 150 Myr... based on ...? what?
"""
Similar to Cummings et al. (2016), Cummings \& Kalirai (2018),
and Cummings et al. (2018) (see also Fritzewski et al. 2020 and Healy
\&
McCullough 2020)
"""

% \subsection{Ways of doing this at different ages}
% Some final discussion is warranted on the applicability of our
% approach more generally.  Given a kinematically
% identified population, this study uses photometry (Gaia HR diagrams;
% TESS rotation) and spectroscopy (GALAH and Gaia-ESO lithium) to
% confirm youth in the stars.  Other approaches are also possible.
% 
% One can use photometric criteria (e.g., Gaia HR diagram of
% everything within some distance; Zari+18), or spectroscopic
% criteria (CITE Berger+18), or with combinations thereof (e.g., CITE
% Zerjal 2019, 2020, Zhou's work).  
% It might even be possible to start using chemical abundance signatures
% (CITE, GALAHDR3?).
% 
% These paths might in certain cases succeed in identifying more
% complete samples of dispersed members of young populations, since they
% do not require kinematic proximity.

@@

The combination of astrometry from Gaia, photometry from TESS, and
spectroscopy from GALAH and Gaia-ESO has clarified a few things about
the structure of \cn.
\begin{itemize}
  %
  \item {\it Over-densities observed in the Gaia 3-D positions and 2-D
    velocities revealed a halo of stars spanning $\approx$500 parsecs
    in the plane of the galaxy.} The earliest reference to this halo
    that we can find is by \citet{kounkel_untangling_2019}.  The halo
    is more precisely described as a leading and trailing tail
    (Figure~\ref{fig:XYZ}), with the leading edge angled toward the
    galactic center, relative to the cluster's orbit in the Galaxy.
    This is consistent with the direction and amplitude of the Milky
    Way's differential rotation
    ($\approx$-0.2\,deg\,Myr$^{-1}$\,kpc$^{-1}$), and similar
    halos/tails have been observed around $\approx$10 other nearby
    open clusters \citep{meingast_2021}.
	%
  \item {\it Isochronal, rotational, and lithium age-dating show that
    the halo of \cn\ is coeval with its core.} In other words, all
    the data are consistent with the halo being real.  The Gaia EDR3
    photometry and astrometry show a main-sequence turnoff consistent
    with an isochronal age of 150\,Myr for both the core and halo
    (Figure~\ref{fig:hr}).  The faintest M dwarfs in the core and halo
    are also brighter than comparison field stars of the same color.
    Gyrochronally, stars in \cn\ spanning $0.5<(Bp-Rp)_0<1.2$
    (spectral types F2V-K3V) overlap with the slow sequence of the
    Pleaides (Figure~\ref{fig:rot}).  At slightly redder
    colors, from $1.2<(Bp-Rp)_0<1.7$ (spectral types K3V-K6V,
    $M_\star$$\approx$ 0.6-0.7\,$M_\odot$), there is a larger
    dispersion in rotation period.  The upper envelope of the rotation
    period distribution at these redder colors extends to
    longer rotation periods than in the Pleiades (11 {\it
    vs.}\ 8 day rotation periods).
    This also supports our {\bf age determination of $XXX$\,Myr for
    the cluster}.
  %
  \item {\it Bonafide members exist out to $\approx$250\,pc in
    separation from the cluster center, and out to $\approx$5\kms\ in
    velocity separation.} Figure~\ref{fig:physical_x_rotn} shows the
    overlap between kinematic and rotational cluster members used to
    make this assessment.
  %
  \item {\it The field star contamination rate increases at larger
    physical and velocity separations from the cluster core.} The
    right panel of Figure~\ref{fig:physical_x_rotn} quantifies the
    statement: $\approx$75\% of the ``inner-most'' members show
    rotation periods consistent with the age of the cluster.  This
    fraction decreases to $\approx$60\% at physical separations of
    $\approx$200\,pc.  Given the number density ratios of cluster to
    field stars at these wide separations ($\approx 10^{-3}$,
    \citealt{meingast_2021}), the implied contamination fractions are
    not as bad as one might fear.
\end{itemize}

Broadly speaking, the existence of the halo is perhaps not very
surprising.  A star with a 1\kms\ velocity difference from the cluster
average will move away from the center at a rate of 1\,pc\,Myr$^{-1}$.
For a $\sim$100\,Myr old cluster, a halo with characteristic size
$\sim$100\,pc is therefore expected, given that the typical velocity
dispersions of open clusters are of the order of kilometers per
second.

On the other hand, our ability to reliably identify members of these
halos is rather surprising.  These stars have
remained hidden in the background of the Galaxy throughout centuries
of modern astronomy.  The consequences that these halo stars will have
in expanding our understanding of planetary and stellar origins are
only just beginning to be understood.
